---
title: "p8105_hw1_ez2384"
output: github_document
date: "2024-09-21"
---

#### **Problem 0.1, 0.2, and loading relevant packages**

*LINK TO DEPOSITORY: <https://github.com/Xs-761/p8105_hw1_ez2384.git>*

```{r message=FALSE}
# Load Prerequisite Packages
  library("ggplot2")
  library("tidyverse")
```

#### **Problem 1**

Load the PENGUINS dataset
```{r}
  data("penguins", package = "palmerpenguins")
```

Description of the PENGUINS dataset:
```{r, results="hide"}
  ?penguins
```

* 8 Variables:
  * SPECIES (a factor denoting penguin species)
  * ISLAND (a factor denoting island in Palmer Archipelago, Antarctica)
  * BILL_LENGTH_MM (a number denoting bill length (millimeters))
  * BILL_DEPTH_MM (a number denoting bill depth (millimeters))
  * FLIPPER_LENGTH_MM (an integer denoting flipper length (millimeters))
  * BODY_MASS_G (an integer denoting body mass (grams))
  * SEX (a factor denoting penguin sex (female, male))
  * YEAR (an integer denoting the study year (2007, 2008, or 2009))
* Size of the Dataset: 344 x 8
```{r}
        nrow(penguins)
        ncol(penguins)
```
  - Mean flipper length
```{r}
        mean(penguins$flipper_length_mm, na.rm=TRUE)
```
  - Scatterplot
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm)) + geom_point(aes(color = species), na.rm=TRUE, size=1) + theme_minimal() + ggtitle("Scatterplot of Bill Length vs. Flipper Length") + theme(plot.title = element_text(hjust = 0.35))
```

#### **Problem 2**

Create the Desired Dataframe
```{r}
  set.seed(2048)      # Select a pseudo random seed for replication purposes
  DF <- data.frame(
          random_sample = rnorm(10),
          char_vector   = letters[1:10],
          factor_vector = factor(sample(c("LEVEL1", "LEVEL2", "LEVEL3"), size = 10, replace = TRUE))  # To randomly select from the 3 levels
  )
  DF$logical_vector <- DF$random_sample>0
```
  
Compute the Means for Each Column and Discuss whether Applicable
```{r}
  mean(DF[["random_sample"]])   # Mean() works for numeric data column
  mean(DF[["logical_vector"]])  # Mean() works for logical data column, treating TRUE as 1 and FALSE as 0
  mean(DF[["char_vector"]])     # Mean() does NOT work for character data types, returning NA
  mean(DF[["factor_vector"]])   # Mean() does NOT work for factor data column, returning NA
```

Applying as.numeric() to the logical, character, and factor variables. Discuss what happens and the reasoning.
```{r}
  conv_num_vector = as.numeric(DF$logical_vector)  # conversion successful: treating TRUE as 1 and FALSE as 0
  conv_num_char   = as.numeric(DF$char_vector)     # warning appears, then returns NA throughout
  conv_num_factor = as.numeric(DF$factor_vector)   # conversion successful: treating the LEVEL CODE of the factor as 1, 2, 3 for the three levels specified
```
